---
title: "proxReg"
author: "Christopher T. Lee"
date: "11/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Running analysis

Download Supplementary Table 1 and place in working directory before proceeding.
```{r eval = F}
newpeaklist = gdata::read.xls("./Supplementary Table 1.xlsx", stringsAsFactors=F)
```

##### To download all the ENCODE ChIP-seq experiments used, follow this procedure:
```{r eval = F}
write.table(sprintf("%s.gz",newpeaklist$Dataset), "./ENCODE_datasets.txt", quote=F, row.names = F, col.names = F)
```

Run the following on bash in your working directory to download the ChIP-seq experiments:

Use `brew install wget` if wget is not already installed.
```
awk '{print "wget http://hgdownload.cse.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeAwgTfbsUniform/"$1}' <(cut -f 1 ENCODE_datasets.txt | grep narrowPeak) > tfbs_wget.sh

bash tfbs_wget.sh
```

### Running proxReg on all data:
```{r eval = F}


```

Heatmaps for three groups
```{r eval = F}
group1 = read.table("~/Desktop/RESEARCH/group1.txt", header =T)
group2 = read.table("~/Desktop/RESEARCH/group2.txt", header =T)
group3 = read.table("~/Desktop/RESEARCH/group3.txt", header =T)

colfunc = colorRampPalette(c("cyan","black","red"))


image(t(as.matrix(group1[,-1])), col = colfunc(10))
image(t(as.matrix(group2[,-1])), col = colfunc(10))
image(t(as.matrix(group3[,-1])), col = colfunc(10))

```





