---
title: "proxReg"
author: "Kai Wang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Table of Contents
* [Figure 2,5: Scatter plot of proxReg vs Poly-Enrich](#figure-2-and-5)
* [Figure 3: Heatmap for 90 proxReg data](#figure-3)
* [Figure 4: Scatter plot of proxReg promoter vs enhancer](#figure-4)

## Figure 2 and 5 
Generate scatter plot of ProxReg results with Poly-Enrich results. Example input file is "Gm12878.Six5.txt", this file contains the FDR and p-values from ProxReg and Poly-Enrich. Users can check the example file for more details.
```{r eval = F}
library(ggplot2)
data   <- read.csv("./datafiles/Gm12878.Six5.txt", header = T, sep = "\t")
data$x <- sign(data$polyenrich.Effect)*-log10(data$polyenrich.pvalue)
data$promoter.y <- -log10(abs(data$promoter.pvalue))*sign(data$promoter.Effect)
data$enhancer.y <- -log10(abs(data$enhancer.pvalue))*sign(data$enhancer.Effect)

#plot promoter results
ggplot(data) +
    geom_point(aes(x = x, y = promoter.y)) +
    theme_classic() +
    geom_vline(xintercept = 0) +
    geom_hline(yintercept = 0) +
    theme(legend.position="none",plot.title = element_text(hjust = 0.5,size = 12),axis.text=element_text(size=10),
          axis.title=element_text(size=10,face="bold")) +
    xlab("Signed -log10 GSE Pvalue") +
    ylab("Signed -log10 ProxReg Pvalue") +
    ggtitle("Example promoter results")

#plot enhancer results
ggplot(data) +
    geom_point(aes(x = x, y = enhancer.y)) +
    theme_classic() +
    geom_vline(xintercept = 0) +
    geom_hline(yintercept = 0) +
    theme(legend.position="none",plot.title = element_text(hjust = 0.5,size = 12),axis.text=element_text(size=10),
          axis.title=element_text(size=10,face="bold")) +
    xlab("Signed -log10 GSE Pvalue") +
    ylab("Signed -log10 ProxReg Pvalue") +
    ggtitle("Example enhancer results")
```


## Figure 3
Heat map for all 90 data sets. The count numbers of points in each quadrant were calculated by in-house python script.  
```{r eval = F}
library(RColorBrewer)
library(gplots)
#read counts data
dat<-read.csv("./datafiles/heat.map.data.csv",header = TRUE,
              row.names=1)
datlog<-log2(dat+2)
d2<-dist(datlog,method = "euclidean",diag = FALSE, upper = TRUE)
c2 <- hclust(d2, method = "ward.D2", members = NULL)

par(cex.main=0.75)
my_palette <- colorRampPalette(c("blue","white","red"))(n = 25)
dm<-as.matrix(datlog)
heatmap.2(dm, 
          col=my_palette, 
          key = TRUE,
          #lhei=c(2,4), lwid=c(2,3.5), keysize=0.75, key.par = list(cex=0.5),
          Rowv=as.dendrogram(c2),
          Colv=FALSE, 
          trace="none",
          key.xlab = "log(count+2)",
          density.info = c("none"),#symkey = FALSE,cexRow = 0.8, cexCol = 1.0,
          #lhei=c(0.3,8),
          margins = c(3,8))
```

## Figure 4
Generate scatter plot of ProxReg promoter results and enhancer results. 
```{r eval = F}
library(ggplot2)
#read data
data   <- read.csv("./datafiles/Gm12878.Six5.txt", header = T, sep = "\t")
data$promoter <- -log10(abs(data$promoter.pvalue))*sign(data$promoter.Effect)
data$enhancer <- -log10(abs(data$enhancer.pvalue))*sign(data$enhancer.Effect)

# 
data["P.value.fdr"]<-p.adjust(data$polyenrich.pvalue, method = "BH")
data$group<-1
data$group[abs(data$P.value.fdr)>=0.05]<-2

#generate scatter plot of ProxReg promoter results and enhancer results
ggplot(data)+
  geom_point(data=subset(data,group==1),aes(x=promoter,y=enhancer),color=c("black")) +
  geom_smooth(data=subset(data,group==1),aes(x=promoter,y=enhancer),method=lm) +
  theme_classic() +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  theme(legend.position="none",plot.title = element_text(hjust = 0.5,size = 12),axis.text=element_text(size=10),
        axis.title=element_text(size=10,face="bold")) +
  xlab("Signed -log10 ProxReg Pvalue (Promoter)") +
  ylab("Signed -log10 ProxReg Pvalue (Enhancer)") +
  ggtitle("Example figure of promoter/enhancer scatter plot")

```

Obtain GO terms that our 35 TFs were assigned to in the human annotation Bioconductor package org.Hs.eg.db, we used Six5 as an example here.
```{r}
# we used the Entrezid to obtain the GO term that assigned to Six5
library(AnnotationDbi)
library(org.Hs.eg.db)
TF2GO = AnnotationDbi::select(org.Hs.eg.db,
                              keys = "147912",
                              keytype = "ENTREZID",
                              columns = c("SYMBOL", "GOALL", "ONTOLOGYALL"))
dat<-TF2GO[TF2GO$ONTOLOGYALL=="BP",]
dat$EVIDENCEALL<-NULL
dat<-unique(dat)
```
