---
title: "proxReg-SuppleFigs"
author: "Christopher T. Lee"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Table of Contents
* [Supplementary Figure 1](#supplementary-figure-1)
	+ [Making SF1](#making-sf1)
* [Supplementary Figure 2](#supplementary-figure-2)
	+ [Making SF2](#making-sf2)

## Supplementary Figure 1

This is to run one iteration of randomizations.
The different randomizations are "unif" or "bylength"

10 trials was run for each peak file.
```{r eval = F}
peak = "./wgEncodeAwgTfbsHaibGm12878Atf3Pcr1xUniPk.narrowPeak"
results_randunif = proxReg(peaks = peak, 
						out_name = NULL, out_path = NULL, genome = "hg19",
						   reglocation = "enhancer", genesets = "GOBP", 
							  randomization = "unif", qc_plots = F)

```

### Making SF1
This makes the figure after aggregating all the data.
```{r eval = F}
newpeaklist = gdata::read.xls("./Supplementary Table 1.xlsx", stringsAsFactors=F)
n = 90
thresh = 0.001
rndmat = matrix(0,n,4)
for (index in (1:n)) {
	rnds = read.table(sprintf("./proxReg_rnd/rndsgene.%s.txt",newpeaklist$V1[index]), header=T)
	rndse = read.table(sprintf("./proxReg_rnd/rndsenh5.%s.txt",newpeaklist$V1[index]), header=T)
	rndmat[index,1:2] = colSums(rnds < thresh)/nrow(rnds)
	rndmat[index,3:4] = colSums(rndse < thresh)/nrow(rndse)

}
boxplot(rndmat[,1:4], main = "alpha = 0.001 \n       Dist to TSS                                Dist to Enhancer", 
		ylim = c(0,0.01), names = c("Unif","By Locus Length","Unif","By Avg DEnh"),
		ylab = "Proportion significant")
abline(h=thresh, col = "red")
abline(v = 2.5, col = "grey", pch = "-")
```



## Supplementary Figure 2

Artificially creates close peaks after randomizing 
```{r eval = F}
devtools::load_all("~/GitHub/chipenrich/") # Need the hidden functions
gs_path = system.file('extdata','vignette_genesets.txt', package='chipenrich')
results_randlength = proxReg("./wgEncodeAwgTfbsHaibGm12878Six5Pcr1xUniPk.narrowPeak", 
						  out_name = NULL, out_path = NULL, genome = "hg19",
					   reglocation = "tss", genesets = gs_path, 
						  randomization = "bylength", qc_plots = F)
peaks_null = results_randlength$peaks

library(chipenrich.data)
ldef = chipenrich.data::locusdef.hg19.nearest_tss@dframe

simulate_peaks_full_random <- function(npeaks, ldef) {
	peaks = matrix(0, npeaks, 3)
	chrs = names(table(ldef$chr))[1:22]
	chr_sim = sample(chrs, npeaks, replace = T, prob = sapply(chrs, function(x){max(ldef$end[ldef$chr==x])}))
	ends_max = sapply(chr_sim, function(chr){max(ldef$end[ldef$chr==chr])})
	ends_sim = sapply(ends_max, function(end){sample(3:end,1)})
	starts_sim = ends_sim -2
	out = data.frame(chr = chr_sim, start = starts_sim, end = ends_sim, stringsAsFactors = F)
	out = out[order(out$chr,out$start),]
	return(out)
}

largego = "GO:0051049" #1717 genes
medgo = "GO:0043410" #471 genes
smallgo = "GO:0070534" #42 genes 


rpeaks = function(nsamples, d0) {
	return(sample(c(-1,1),nsamples,replace = T)*round(d0*rexp(nsamples)))
}

add_simulated_peaks_near_genes <- function(sim_peaks, goid, proppeaks, d0) {
	#geneset.list = read.table("/espresso/share/reverse_query_initial_GO_2.txt",header = T, stringsAsFactors = F)
	geneset.list = read.table("./reverse_query_initial_GO_2.txt",header = T, stringsAsFactors = F) 
    go_genes = strsplit(geneset.list$Geneset.Genes[geneset.list$GenesetID==goid], split=",")[[1]]

    simgenes = sample(intersect(go_genes,names(table(sim_peaks$gene_id))),
    				  floor(nrow(sim_peaks)*proppeaks), replace = TRUE)
    simindices = do.call(rbind, 
    					 lapply(simgenes, function(x) {sim_peaks[sim_peaks$gene_id==x,][1,]}))
    simindices$dist_to_tss = rpeaks(nrow(simindices), d0)

    out = rbind(sim_peaks, simindices)
	return(out)
}

single_proxReg = function(go_id, geneset, peaks, regloc) {

	# Genes in the geneset
	go_genes = geneset@set.gene[[go_id]]
	
	# Background genes and the background presence of a peak
	b_genes = peaks$gene_id %in% go_genes
	if (regloc == "tss"){
		x = peaks$scaled_dtss[b_genes]
		y = peaks$scaled_dtss[!b_genes]
	}
	if (regloc == "enhancer"){
		x = peaks$scaled_denh[b_genes]
		y = peaks$scaled_denh[!b_genes]
	}
	x <- x[is.finite(x)]
	y <- y[is.finite(y)]
	
	# Information about the geneset
	r_go_id = go_id
	r_go_genes_num = length(go_genes)

	# Information about peak genes
	go_genes_peak = peaks$gene_id[b_genes]
	r_go_genes_peak = length(table(go_genes_peak))
	r_go_genes_peak_num = length(go_genes_peak)

	r_effect = NA
	r_pval = NA
	
	if (length(x)== 0L | length(y) == 0L) {
		sprintf("Geneset: %s has zero peaks in one group. NAs given", go_id)
	} else {
		tryCatch(
			{ #Code from wilcox.test.default
			r <- rank(c(x, y))
			n.x <- as.double(length(x))
			n.y <- as.double(length(y))
			STATISTIC <- c(W = sum(r[seq_along(x)]) - n.x * (n.x + 1)/2)
			NTIES <- table(r)
			z <- STATISTIC - n.x * n.y/2
			SIGMA <- sqrt((n.x * n.y/12) * ((n.x + n.y + 1) - 
							sum(NTIES^3 - NTIES)/((n.x + n.y) * (n.x + n.y - 
								1))))
			z = z/SIGMA
			
			# Results from the Wilcoxon test
			r_effect = -z; #We want closer to be positive effect
			r_pval = 2 * min(pnorm(z), 
							 pnorm(z, lower.tail = FALSE))
			},
			error = {function(e) {warning(
				sprintf("Error in geneset: %s. NAs given", go_id))
			}}
		)
	}
	
	out = data.frame(
		"P.value"=r_pval,
		"Geneset ID"=r_go_id,
		"N Geneset Genes"=r_go_genes_num,
		"Geneset Peak Genes"=r_go_genes_peak,
		"N Geneset Peak Genes"=r_go_genes_peak_num,
		"Effect"=r_effect,
		stringsAsFactors = FALSE)
	
	return(out)
}

```

```{r eval = F}
ldef_list = setup_locusdef(ldef_code = "nearest_tss", "hg19", randomization = NULL)
ldef = ldef_list[['ldef']]
tss = ldef_list[['tss']]
test_power = function(index, go_id, proppeaks, d0) {
	print(index)
	
	peakobj = read_bed(sprintf("~/Desktop/RESEARCH/ENCODEpeaks/%s",newpeaklist$V1[index]))
	message("Assigning peaks to genes with assign_peaks(...) ..")
	assigned_peaks = assign_peaks(peakobj, ldef, tss)

	assigned_peaks$log_dtss = log(abs(assigned_peaks$dist_to_tss)+1)
	assigned_peaks$log_gene_ll = log(assigned_peaks$gene_locus_end-assigned_peaks$gene_locus_start)
	pred_log_dtss = as.numeric(mgcv::predict.gam(chipenrich.data::spline.log_dtss.90ENCODE, assigned_peaks, type="link"))
	assigned_peaks$scaled_dtss = assigned_peaks$log_dtss-pred_log_dtss
	message("Randomizing by picking gene ids by locus length...")
	peakstemp = assigned_peaks[,c("peak_id","gene_id","gene_locus_start","gene_locus_end")]
	peakstemp$gene_locus_length = peakstemp$gene_locus_end-peakstemp$gene_locus_start
	peakstemp$dupe_genes = duplicated(peakstemp$gene_id)
	peakstemp = peakstemp[sample(1:nrow(peakstemp)),]
	peakstemp2 = peakstemp[!peakstemp$dupe_genes,]
	peakstemp2 = peakstemp2[order(peakstemp2$gene_locus_length),]
	rownames(peakstemp2) = 1:nrow(peakstemp2)
	peakstemp2$group = floor((as.numeric(rownames(peakstemp2))+99)/100)
	peakstemp$group = sapply(peakstemp$gene_id, function(x){peakstemp2$group[peakstemp2$gene_id==x]})
	peakstemp$gene_id = sapply(peakstemp$group, function(x){sample(peakstemp2$gene_id[peakstemp2$group==x],1)})
	assigned_peaks$gene_id_pre = assigned_peaks$gene_id
	assigned_peaks = merge(assigned_peaks[,-5], peakstemp[,c("peak_id","gene_id")], by = "peak_id")
	peaks_null = assigned_peaks
	npeaks = nrow(peaks_null)
	assigned_peaks = add_simulated_peaks_near_genes(peaks_null,go_id,proppeaks,
													 d0 = d0 )
	assigned_peaks$log_dtss = log(abs(assigned_peaks$dist_to_tss)+1)
	assigned_peaks$log_gene_ll = log(assigned_peaks$gene_locus_end-assigned_peaks$gene_locus_start)
	pred_log_dtss = as.numeric(mgcv::predict.gam(chipenrich.data::spline.log_dtss.90ENCODE, assigned_peaks, type="link"))
	assigned_peaks$scaled_dtss = assigned_peaks$log_dtss-pred_log_dtss
	
	out = single_proxReg(go_id, chipenrich.data::geneset.GOBP.hsa, assigned_peaks, regloc = "tss")
}

#prox_power = lapply(1:90, function(x) {test_power(x, medgo, 0.01, 500)})
#prox_power = do.call(rbind, prox_power)

gos = c(medgo, largego)
d0s = c(100,500,1000)
propadd = c(0.001,0.005,0.01)
```

```{r eval = F}
ntrials = 100
#Datasets : 65, 34, 32
index =34
indicies = rep(index, 200)
prox_power_array = array(list(),c(2,3,3))
for (i in 1:length(gos)) {
	for (j in 1:length(propadd)) {
		for (k in 1:length(d0s)) {
			prox_power = lapply(indicies, function(x) {test_power(x, gos[i], propadd[j],d0s[k] )})
			prox_power = do.call(rbind, prox_power)
			prox_power_array[i,j,k][[1]] = prox_power
		}
	}
}

prox_powers = apply(prox_power_array,c(1,2,3), function(x){mean(x[[1]]$P.value<0.05)})

#saveRDS(prox_powers, "~/Desktop/RESEARCH/prox_powers.rds")
```


### Making SF2

Preprocessed data is supplied on Github
```{r eval = F}
prox_powers = readRDS("./prox_powers.rds")
#Supplementary Figure 2
plot(NULL,xlim = c(0,1), ylim = c(0,100), xlab = "% peaks added", ylab = "% significant", main = "Power")
points(propadd*100, prox_powers[1,,1]*100 ,type ="b", col = "orange")
points(propadd*100, prox_powers[1,,2]*100 ,type ="b", col = "red")
points(propadd*100, prox_powers[1,,3]*100 ,type ="b", col = "red4")

points(propadd*100, prox_powers[2,,1]*100 ,type ="b", col = "violet")
points(propadd*100, prox_powers[2,,2]*100 ,type ="b", col = "blue")
points(propadd*100, prox_powers[2,,3]*100 ,type ="b", col = "blue4")

legend("bottomright",legend = c("SmallGS, d0 = 100","SmallGS, d0 = 500","SmallGS, d0 = 1000",
								"LargeGS, d0 = 100","LargeGS, d0 = 500","LargeGS, d0 = 1000"), pch =  15,
	   col = c("orange","red","red4","violet","blue","blue4"))
```

